---
title: "Liberty Lynx Game One"
output: github_document
date: "2024-10-12"
prefer-html: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE
                      ,warning = FALSE)
```

## What a start!

This is a little R investigation to the question:

> How did the Lynx come back to win so decisively?

Let's get started with some WNBA R packages.

```{r load-pbp}
library(wehoop)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggforce)
library(gtExtras)
wnba_pbp <- wehoop::load_wnba_pbp()

```

```{r game-data}
finals_game_one_id <- 401726988

finals_game_one_pbp <- wnba_pbp %>%
  filter(game_id == finals_game_one_id)

```


```{r gameone-analysis}
#finals_game_one_pbp %>%
#  View()
# 
player_court_id <- 2987891
player_phee_id <- 3917450
# 
# finals_game_one_pbp %>%
#   filter(athlete_id_1 == player_court_id) %>%
#   filter(shooting_play == TRUE) %>%
#   View()
# 
# finals_game_one_pbp %>%
#   filter(athlete_id_1 == player_phee_id) %>%
#   filter(shooting_play == TRUE) %>%
#   View()
# 

```

```{r court-shot-chart}

finals_game_one_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  filter(athlete_id_1 == player_court_id) %>%
  filter(shooting_play == TRUE) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw, color = scoring_play, size = 8)) +
  geom_jitter() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle("This chart was an excuse to watch Courtney Williams replays") +
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5))

```

```{r phee-shot-chart}

finals_game_one_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  filter(athlete_id_1 == player_phee_id) %>%
  filter(shooting_play == TRUE) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw, color = scoring_play, size = 8)) +
  geom_jitter() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle("Phee is incredible") +
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5))

```

```{r cars}

# find all 4 point_plays

wnba_pbp %>%
  mutate(day = as.Date(wallclock)) %>%
  group_by(home_team_mascot, away_team_mascot, season, game_id, day, athlete_id_1,period_display_value, clock_display_value) %>%
  summarize(total_points = sum(score_value)) %>%
  filter(total_points > 3) %>%
  arrange(desc(total_points))
  
```


There was actually one FIVE-point play - kelsey mitchell was fouled by arike shooting a 3 on Sept 1st. Arike got a technical for complaining - so it was a 5-point play!


### Four point games in 2024... wait, what?

```{r five-point-play}
wnba_pbp %>%
  mutate(day = as.Date(wallclock)) %>%
  group_by(home_team_mascot, away_team_mascot, season, game_id, day, athlete_id_1,period_display_value, clock_display_value) %>%
  summarize(total_points = sum(score_value)) %>%
  filter(total_points > 3) %>%
  arrange(desc(total_points)) %>%
  select(total_points) %>%
  group_by(total_points) %>%
  summarise(plays_in_2024 = n()) %>%
  rename(total_points_on_play = total_points) %>%
  select(plays_in_2024, total_points_on_play) %>%
  knitr::kable()
#%>%
 # gt() %>% 
  #gt_theme_guardian() %>% 
  #tab_header(title = "Four point games in 2024... wait, what?")

```
```{r deep dive}
finals_game_one_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  filter(athlete_id_1 == player_court_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(sequence_number == 482) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw,  size = 8)) +
  geom_point() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle("Wasn't this shot from the edge of the key?") +
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5))


```


Now here's Phee in the post-season

```{r}

wnba_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  mutate(day = as.Date(wallclock)) %>%
  filter(day >= "2024-09-22") %>%
  filter(athlete_id_1 == player_court_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(!grepl("Free Throw", type_text)) %>%
  filter(grepl("Jump Shot", type_text)) %>%
  filter(scoring_play == TRUE) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw, color = type_text, size = 8, alpha = 0.4)) +
  geom_jitter() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle(" Courtney Williams is KILLING IT with high-momentum jump shots near the free-throw line") +
    theme_void() +
   # theme(plot.title = element_text(hjust = 0.5)) +
    theme(legend.text=element_text(size=6))

```
```{r}

wnba_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  mutate(day = as.Date(wallclock)) %>%
  filter(day >= "2024-09-22") %>%
  filter(athlete_id_1 == player_court_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(!grepl("Free Throw", type_text)) %>%
  filter(grepl("Jump Shot", type_text)) %>%
  filter(scoring_play == FALSE) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw, color = type_text, size = 8, alpha = 0.4)) +
  geom_jitter() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle(" Her traditional KILL ZONE is less successful this post-season. Here are her misses.") +
    theme_void() +
   # theme(plot.title = element_text(hjust = 0.5)) +
    theme(legend.text=element_text(size=6))

```


### "Since the start of this postseason, Courtney has struggled with pullup jump shots - except when she's mid-sprint!!"

```{r}

wnba_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  mutate(day = as.Date(wallclock)) %>%
  filter(day >= "2024-09-22") %>%
  filter(athlete_id_1 == player_court_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(!grepl("Free Throw", type_text)) %>%
  group_by(type_text, scoring_play ) %>%
  summarize(count = n()) %>%
  pivot_wider(names_from = scoring_play , values_from = count, values_fill = 0) %>%
  mutate(attempts = `FALSE` + `TRUE`) %>%
  mutate(success_rate = `TRUE` / attempts) %>%
  rename(makes = `TRUE`) %>%
  arrange(desc(success_rate)) %>%
  select(-`FALSE`) %>%
  filter(grepl("Jump Shot", type_text)) %>%
  mutate(makes_and_attempts = paste0(makes, " for ", attempts)) %>%
  rename(jumpshot_type = type_text) %>%
  select(jumpshot_type, makes_and_attempts, success_rate)  %>%
  knitr::kable()
#%>%
  #gt() %>% 
  #gt_theme_guardian() %>% 
  #tab_header(title = "Since the start of this postseason, Courtney has struggled with pullup jump shots - except when she's mid-sprint!!")

```


Let's compare with Phee.

### Napheesa Collier shows a similar inclination towards more complex jump shots. But hers are pivot-y!

```{r}

wnba_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  mutate(day = as.Date(wallclock)) %>%
  filter(day >= "2024-09-22") %>%
  filter(athlete_id_1 == player_phee_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(!grepl("Free Throw", type_text)) %>%
  group_by(type_text, scoring_play ) %>%
  summarize(count = n()) %>%
  pivot_wider(names_from = scoring_play , values_from = count, values_fill = 0) %>%
  mutate(attempts = `FALSE` + `TRUE`) %>%
  mutate(success_rate = `TRUE` / attempts) %>%
  rename(makes = `TRUE`) %>%
  arrange(desc(attempts)) %>%
  select(-`FALSE`) %>%
  filter(grepl("Jump Shot", type_text)) %>%
  mutate(makes_and_attempts = paste0(makes, " for ", attempts)) %>%
  rename(jumpshot_type = type_text) %>%
  select(jumpshot_type, makes_and_attempts, success_rate) %>%
  knitr::kable()
#%>%
 # gt() %>% 
  #gt_theme_guardian() %>% 
#  tab_header(title = "Napheesa Collier shows a similar inclination towards more complex jump shots. But hers are pivot-y!")

```


```{r}

wnba_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  mutate(day = as.Date(wallclock)) %>%
  filter(day >= "2024-09-22") %>%
  filter(athlete_id_1 == player_phee_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(!grepl("Free Throw", type_text)) %>%
  filter(grepl("Jump Shot", type_text)) %>%
  filter(scoring_play == TRUE) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw, color = type_text, size = 8, alpha = 0.4)) +
  geom_jitter() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle(" High-pivot Napheesa Collier success this postseason!!") +
    theme_void() +
   # theme(plot.title = element_text(hjust = 0.5)) +
    theme(legend.text=element_text(size=6))

```

```{r}

wnba_pbp %>%
  mutate(coordinate_x_raw = 50-coordinate_x_raw) %>%
  mutate(day = as.Date(wallclock)) %>%
  filter(day >= "2024-09-22") %>%
  filter(athlete_id_1 == player_phee_id) %>%
  filter(shooting_play == TRUE) %>%
  filter(!grepl("Free Throw", type_text)) %>%
  filter(grepl("Jump Shot", type_text)) %>%
  filter(scoring_play == FALSE) %>%
  ggplot(aes(x = coordinate_x_raw, y = coordinate_y_raw, color = type_text, size = 8, alpha = 0.4)) +
  geom_jitter() +
    geom_segment(aes(x = 0, y = -4, xend = 50, yend = -4, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = 13, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 17, y = -4, xend = 17, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 19, y = -4, xend = 19, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 31, y = -4, xend = 31, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 33, y = -4, xend = 33, yend = 13, alpha =0.1), inherit.aes = FALSE) +
    geom_segment(aes(x = 22, y = -0.75, xend = 28, yend = -0.75, alpha =0.1), inherit.aes = FALSE) +
    geom_circle(aes(x0 = 25, y0 = 0, r = 0.75), inherit.aes = FALSE) +
    geom_arc(aes(x0 = 25, y0= 0, r = 22.145, start = -1.2, end = 1.2), inherit.aes = FALSE) +
    #geom_segment(aes(x = 3, y = -4, xend = 3, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    #geom_segment(aes(x = 47, y = -4, xend = 47, yend = 10, alpha =0.1), inherit.aes = FALSE) +
    scale_size(guide = "none") +
    scale_alpha(guide = "none") +
    xlab("")+
    ylab("")+
    ggtitle(" High-pivot Napheesa Collier success this postseason!!") +
    theme_void() +
   # theme(plot.title = element_text(hjust = 0.5)) +
    theme(legend.text=element_text(size=6))

```
